\vspace{-.5cm} Lights: a generalized joint model for high-dimensional multivariate longitudinal data and censored durations \vspace{.5cm}

### 2021-03-18 paper.tex
{\gamma_k^0}^{(w+1)} for the \gamma_k update
Update the proximal gradient descent expression
Add back-tracking line search in ISTA algorithm

\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{mathrsfs} 
\usepackage{natbib}
\usepackage{setspace}
\usepackage{dsfont}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subfigure}

\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{orange}{#1}}
\SetCommentSty{mycommfont}
\SetKwInput{KwInput}{Input} 
\SetKwInput{KwOutput}{Output}

@@

\begin{algorithm}[H]
\DontPrintSemicolon
\caption{Proximal gradient descent ISTA$_{\text{sg} l_1}$($v$, $f$, $\tilde \eta$, $\zeta$)}
\label{alg:ISTA}
\KwInput{A vector $v$, a function $f$ (its gradient), hyper-parameters $\tilde \eta$ and $\zeta$}
\KwOutput{$\flat^{\breve{w}^{max}}$}
Initialize $\flat^{(0)}=v$ \\
\For{$\breve{w} = 0,\ldots,\breve{w}^{max}$}    
{ 
  $\begin{aligned}
  \flat^{(\breve{w}+1)} \in &\argmin_u \Big\{\dfrac{1}{2\varrho_{\breve{w}}} \left\lVert u - \big(\flat^{(\breve{w})} - \varrho_{\breve{w}} \dfrac{\partial f(u)}{\partial u} \big)\right\rVert_2^2 + \zeta \norm{u}_{\text{sg} l_1, \tilde{\eta}} \Big\} \\
  &\coloneqq\text{prox}_{\text{sg} l_1, \tilde{\eta}, \zeta} \Big(\flat^{(\breve{w})} - \varrho_{\breve{w}} \dfrac{\partial f(u)}{\partial u} \Big)
  \end{aligned}$
}
\textbf{Return:} {$\flat^{\breve{w}^{max}}$}
\end{algorithm}

@@

Concerning the update of ${\gamma_k^0}^{(w)}$ and $\gamma_k^{(w)}$ for $k=0, \ldots, K-1$, let us follow the same strategy by denoting
\begin{align*}
  Q^{(w)}_{n,k}(\gamma_k^0, \gamma_k) = -n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ &\delta_i \hat \E_{\theta^{(w)}}[ \log \circ \tilde g_{i, t_i, \gamma_k^0, \gamma_k, \beta_k^{(w+1)}}^1(b_i) | t_i, \delta_i, y_i] \\
  & - \sum_{j=1}^J \lambda_0^{(w)}(t_j) \hat \E_{\theta^{(w)}}[ \tilde g_{i, t_j, \gamma_k^0, \gamma_k, \beta_k^{(w+1)}}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \Big] 
\end{align*}
based on the quantities involved in $\cQ_n(\theta, \theta^{(w)})$ that depend on $\gamma_k^0$ and $\gamma_k$, with $\tilde g_{i, t_j, \gamma_k^0, \gamma_k, \beta_k^{(w+1)}}^1$ defined in~\eqref{eq:g-tilde-1-2}. The update for ${\gamma_k^0}^{(w)}$ therefore requires to solve the following minimization problem
\begin{equation}
  \label{eq:minimization-gamma_k^0}
  {\gamma_k^0}^{(w+1)} \in \argmin_{\gamma_k^0 \in \R^p} Q^{(w)}_{n,k}(\gamma_k^0, {\gamma_k}^{(w)}) + \zeta_{1,k} \norm{\gamma_k^0}_{\text{en}, \eta}.
\end{equation}
We then rewrite problem~\eqref{eq:minimization-gamma_k^0} as the following differentiable problem with box constraints
\begin{equation}
  \label{eq:sub-problem-gamma_k^0}
  \begin{split}
    \text{minimize}& \quad \quad Q_{n, k}^{(w)}({\gamma_k^0}^+ - {\gamma_k^0}^-, {\gamma_k}^{(w)}) + \zeta_{1,k} \big((1 - \eta) \sum_{j=1}^p ({\gamma_{k,j}^0}^+ + {\gamma_{k,j}^0}^-) + \dfrac \eta 2 \norm{{\gamma_k^0}^+ - {\gamma_k^0}^-}_2^2 \big) \\
    \text{subject to}& \quad \quad {\gamma_{k,j}^0}^+ \geq 0 \text{ and } {\gamma_{k,j}^0}^- \geq 0 \text{ for } j = 1, \ldots, p.
  \end{split} 
\end{equation}
The gradient is here given by
\begin{align}
  \label{eq:grad-Q-1}
  \frac{\partial Q_{n,k}^{(w)}(\gamma_k^0, {\gamma_k}^{(w)})}{\partial \gamma_k^0} = -n^{-1} &\sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ \delta_i - \sum_{j=1}^J \lambda_0^{(w)}(t_j) \nonumber \\
  &\times \hat \E_{\theta^{(w)}}[ \tilde g_{i, t_j, \gamma_k^0, {\gamma_k}^{(w)}, \beta_k^{(w+1)}}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \Big] x_i.
\end{align}
Then, the update for $\gamma_k^{(w)}$ requires to solve the following minimization problem
\begin{equation}
  \label{eq:minimization-gamma_k}
  \gamma_k^{(w+1)} \in \argmin_{\gamma_k \in \R^{L\mathcal{A}}} Q^{(w)}_{n,k}({\gamma_k^0}^{(w)}, \gamma_k) + \zeta_{3,k} \norm{\gamma_k}_{\text{sg} l_1, \tilde{\eta}}.
\end{equation}
We treat problem~\eqref{eq:minimization-gamma_k} in the same way as~\eqref{eq:minimization-beta_k}, that is
\begin{equation}
  \label{eq:gamma-update}
  \gamma_k^{(w+1)} = \text{ISTA}_{\text{sg} l_1}\big(\gamma_k^{(w)}, Q^{(w)}_{n,k}({\gamma_k^0}^{(w)}, \cdot), \tilde \eta, \zeta_{3,k}\big)
\end{equation}
with the gradient being known since
\begin{align}
  \label{eq:grad-Q-2}
  \frac{\partial Q_{n,k}^{(w)}({\gamma_k^0}^{(w)}, \gamma_k)}{\partial \gamma_{k,a}^l} &= -n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ \delta_i  \hat \E_{\theta^{(w)}}[ \varphi_a(t_i, {\beta_k^l}^{(w+1)}, b_i^l) | t_i, \delta_i, y_i] \\ \nonumber
   &- \sum_{j=1}^J \lambda_0^{(w)}(t_j) \hat \E_{\theta^{(w)}}[ \varphi_a(t_j, {\beta_k^l}^{(w+1)}, b_i^l) \tilde g_{i, t_j, {\gamma_k^0}^{(w)}, \gamma_k, \beta_k^{(w+1)}}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \Big]
\end{align}
for all $a = 1, \ldots, \mathcal{A}$ and $l = 1, \ldots, L$.

>>>>

\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{mathrsfs} 
\usepackage{natbib}
\usepackage{setspace}
\usepackage{dsfont}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subfigure}
\usepackage{algpseudocode}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{orange}{#1}}
\SetCommentSty{mycommfont}
\SetKwInput{KwInput}{Input} 
\SetKwInput{KwOutput}{Output}

@@

\begin{algorithm}[H]
\DontPrintSemicolon
\caption{Proximal gradient descent ISTA$_{\text{sg} l_1}$($v$, $f$, $\tilde \eta$, $\zeta$)}
\label{alg:ISTA}
\KwInput{A vector $v$, a function $f$ (its gradient), hyper-parameters $\tilde \eta$ and $\zeta$}
\KwOutput{$\flat^{\breve{w}^{max}}$}
Define $p_{\varrho}(y) = \argmin_u \Big\{\dfrac{1}{2\varrho} \left\lVert u - \big(y - \varrho \dfrac{\partial f}{\partial u}(y) \big)\right\rVert_2^2 + \zeta \norm{u}_{\text{sg} l_1, \tilde{\eta}} \Big\}$\\
Initialize $\flat^{(0)}=v$, $\varrho_0 > 0$, some $\epsilon > 1$ \\
\For{$\breve{w} = 0,\ldots,\breve{w}^{max}$}    
{   Find the smallest nonnegative integer $i^{\breve{w}}$ such that with $\tilde{\varrho}_{\breve{w}} = \epsilon^{i^{\breve{w}}}\varrho_{\breve{w}}$
	$f(p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})})) \leq f(\flat^{(\breve{w})}) + \big<\big(p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})}) - \flat^{(\breve{w})}\big), \dfrac{\partial f}{\partial u}(\flat^{(\breve{w})})\big> + \dfrac{1}{2\tilde{\varrho}_{\breve{w}}}\norm{p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})}) - \flat^{(\breve{w})}}_2^2$\\
  Set $\varrho_{(\breve{w} + 1)} = \tilde{\epsilon}^{i^{\breve{w}}}  \varrho_{\breve{w}}$
  $\begin{aligned}
  \flat^{(\breve{w}+1)} \in &\argmin_u \Big\{\dfrac{1}{2\varrho_{(\breve{w} + 1)}} \left\lVert u - \big(\flat^{(\breve{w})} - \varrho_{(\breve{w} + 1)} \dfrac{\partial f}{\partial u}(\flat^{(\breve{w})}) \big)\right\rVert_2^2 + \zeta \norm{u}_{\text{sg} l_1, \tilde{\eta}} \Big\} \\
  &\coloneqq\text{prox}_{\text{sg} l_1, \tilde{\eta}, \zeta} \Big(\flat^{(\breve{w})} - \varrho_{(\breve{w} + 1)} \dfrac{\partial f}{\partial u}(\flat^{(\breve{w})}) \Big)
  \end{aligned}$
}
\textbf{Return:} {$\flat^{\breve{w}^{max}}$}
\end{algorithm}

@@

Concerning the update of ${\gamma_k^0}^{(w)}$ and $\gamma_k^{(w)}$ for $k=0, \ldots, K-1$, let us follow the same strategy by denoting
\begin{align*}
  Q^{(w)}_{n,k}(\gamma_k^0, \gamma_k) = -n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ &\delta_i \hat \E_{\theta^{(w)}}[ \log \circ \tilde g_{i, t_i, \gamma_k^0, \gamma_k, \beta_k^{(w+1)}}^1(b_i) | t_i, \delta_i, y_i] \\
  & - \sum_{j=1}^J \lambda_0^{(w)}(t_j) \hat \E_{\theta^{(w)}}[ \tilde g_{i, t_j, \gamma_k^0, \gamma_k, \beta_k^{(w+1)}}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \Big] 
\end{align*}
based on the quantities involved in $\cQ_n(\theta, \theta^{(w)})$ that depend on $\gamma_k^0$ and $\gamma_k$, with $\tilde g_{i, t_j, \gamma_k^0, \gamma_k, \beta_k^{(w+1)}}^1$ defined in~\eqref{eq:g-tilde-1-2}. The update for ${\gamma_k^0}^{(w)}$ therefore requires to solve the following minimization problem
\begin{equation}
  \label{eq:minimization-gamma_k^0}
  {\gamma_k^0}^{(w+1)} \in \argmin_{\gamma_k^0 \in \R^p} Q^{(w)}_{n,k}(\gamma_k^0, {\gamma_k}^{(w)}) + \zeta_{1,k} \norm{\gamma_k^0}_{\text{en}, \eta}.
\end{equation}
We then rewrite problem~\eqref{eq:minimization-gamma_k^0} as the following differentiable problem with box constraints
\begin{equation}
  \label{eq:sub-problem-gamma_k^0}
  \begin{split}
    \text{minimize}& \quad \quad Q_{n, k}^{(w)}({\gamma_k^0}^+ - {\gamma_k^0}^-, {\gamma_k}^{(w)}) + \zeta_{1,k} \big((1 - \eta) \sum_{j=1}^p ({\gamma_{k,j}^0}^+ + {\gamma_{k,j}^0}^-) + \dfrac \eta 2 \norm{{\gamma_k^0}^+ - {\gamma_k^0}^-}_2^2 \big) \\
    \text{subject to}& \quad \quad {\gamma_{k,j}^0}^+ \geq 0 \text{ and } {\gamma_{k,j}^0}^- \geq 0 \text{ for } j = 1, \ldots, p.
  \end{split} 
\end{equation}
The gradient is here given by
\begin{align}
  \label{eq:grad-Q-1}
  \frac{\partial Q_{n,k}^{(w)}(\gamma_k^0, {\gamma_k}^{(w)})}{\partial \gamma_k^0} = -n^{-1} &\sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ \delta_i - \sum_{j=1}^J \lambda_0^{(w)}(t_j) \nonumber \\
  &\times \hat \E_{\theta^{(w)}}[ \tilde g_{i, t_j, \gamma_k^0, {\gamma_k}^{(w)}, \beta_k^{(w+1)}}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \Big] x_i.
\end{align}
Then, the update for $\gamma_k^{(w)}$ requires to solve the following minimization problem
\begin{equation}
  \label{eq:minimization-gamma_k}
  \gamma_k^{(w+1)} \in \argmin_{\gamma_k \in \R^{L\mathcal{A}}} Q^{(w)}_{n,k}({\gamma_k^0}^{(w+1)}, \gamma_k) + \zeta_{3,k} \norm{\gamma_k}_{\text{sg} l_1, \tilde{\eta}}.
\end{equation}
We treat problem~\eqref{eq:minimization-gamma_k} in the same way as~\eqref{eq:minimization-beta_k}, that is
\begin{equation}
  \label{eq:gamma-update}
  \gamma_k^{(w+1)} = \text{ISTA}_{\text{sg} l_1}\big(\gamma_k^{(w)}, Q^{(w)}_{n,k}({\gamma_k^0}^{(w+1)}, \cdot), \tilde \eta, \zeta_{3,k}\big)
\end{equation}
with the gradient being known since
\begin{align}
  \label{eq:grad-Q-2}
  &\frac{\partial Q_{n,k}^{(w)}({\gamma_k^0}^{(w+1)}, \gamma_k)}{\partial \gamma_{k,a}^l} = -n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ \delta_i  \hat \E_{\theta^{(w)}}[ \varphi_a(t_i, {\beta_k^l}^{(w+1)}, b_i^l) | t_i, \delta_i, y_i] \\ \nonumber
   &\qquad\qquad - \sum_{j=1}^J \lambda_0^{(w)}(t_j) \hat \E_{\theta^{(w)}}[ \varphi_a(t_j, {\beta_k^l}^{(w+1)}, b_i^l) \tilde g_{i, t_j, {\gamma_k^0}^{(w+1)}, \gamma_k, \beta_k^{(w+1)}}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \Big]
\end{align}
for all $a = 1, \ldots, \mathcal{A}$ and $l = 1, \ldots, L$.
